---
title: "googleapi4"
author: "Hanna Chang"
format: html
editor: visual
---

```{r}
Sys.setenv(GOOGLE_MAPS_KEY = "")
```

```{r}
library(httr)
library(jsonlite)
library(dplyr)
library(purrr)
library(tidyr)
library(sf)
```

## Turn stops + trip endpoints into sf objects & find nearest stops
```{r}
stops_sf <- stops %>% 
  st_as_sf(coords = c("stop_lon", "stop_lat"),
           crs = 4326,
           remove = FALSE)

starts_sf <- sample_trips %>% 
  st_as_sf(coords = c("start_lon", "start_lat"),
           crs = 4326,
           remove = FALSE)

ends_sf <- sample_trips %>% 
  st_as_sf(coords = c("end_lon", "end_lat"),
           crs = 4326,
           remove = FALSE)

nearest_start_idx <- st_nearest_feature(starts_sf, stops_sf)
nearest_end_idx   <- st_nearest_feature(ends_sf,   stops_sf)

sample_trips_snapped <- sample_trips %>% 
  mutate(
    start_stop_id_snap  = stops_sf$stop_id [nearest_start_idx],
    start_stop_lat_snap = stops_sf$stop_lat[nearest_start_idx],
    start_stop_lon_snap = stops_sf$stop_lon[nearest_start_idx],
    
    end_stop_id_snap    = stops_sf$stop_id [nearest_end_idx],
    end_stop_lat_snap   = stops_sf$stop_lat[nearest_end_idx],
    end_stop_lon_snap   = stops_sf$stop_lon[nearest_end_idx]
  )

sample_trips_snapped %>% 
  select(trip_id,
         start_lat, start_lon,
         start_stop_id_snap, start_stop_lat_snap, start_stop_lon_snap,
         end_lat, end_lon,
         end_stop_id_snap, end_stop_lat_snap, end_stop_lon_snap) %>% 
  head()
```

## Use snapped stops for Google Directions (transit)

### Debug calls
```{r}
debug_google_call <- function(origin_lat, origin_lon,
                              dest_lat, dest_lon,
                              mode = c("bicycling", "transit"),
                              departure_time = NULL,
                              api_key) {

  mode <- match.arg(mode)

  if (is.null(departure_time)) {
    departure_unix <- as.integer(Sys.time())
  } else {
    departure_unix <- as.integer(as.POSIXct(departure_time, tz = "UTC"))
  }

  base_url <- "https://maps.googleapis.com/maps/api/directions/json"

  q <- list(
    origin      = sprintf("%f,%f", origin_lat, origin_lon),
    destination = sprintf("%f,%f", dest_lat,  dest_lon),
    mode        = mode,
    key         = api_key
  )
  if (mode == "transit") q$departure_time <- departure_unix

  full_url <- modify_url(base_url, query = q)
  cat("DEBUG URL:\n", full_url, "\n\n")

  res  <- GET(full_url)
  txt  <- content(res, "text", encoding = "UTF-8")
  cont <- fromJSON(txt)

  status <- cont$status
  err    <- cont$error_message %||% NA_character_

  duration_min <- if (identical(status, "OK")) {
    cont$routes[[1]]$legs[[1]]$duration$value / 60
  } else NA_real_

  tibble(
    duration_min  = duration_min,
    route_summary = NA_character_,
    status        = status,
    error_message = err
  )
}
```

```{r}
test_debug <- debug_google_call(
  origin_lat = 44.9778,   # Downtown Minneapolis
  origin_lon = -93.2650,
  dest_lat   = 44.9733,   # UMN East Bank area
  dest_lon   = -93.2260,
  mode       = "transit",
  departure_time = "2025-06-04 08:00:00",
  api_key    = "()"
)

test_debug
```

## Function redo
```{r}
google_snapped <- sample_trips_snapped %>% 
  slice_head(n = 500) %>%          #  first 500 trips
  rowwise() %>% 
  mutate(
    transit = list(
      debug_google_call(
        origin_lat  = start_stop_lat_snap,
        origin_lon  = start_stop_lon_snap,
        dest_lat    = end_stop_lat_snap,
        dest_lon    = end_stop_lon_snap,
        mode        = "transit",
        departure_time = StartTime,
        api_key     = Sys.getenv("GOOGLE_MAPS_KEY")
      )
    )
  ) %>% 
  unnest(transit)

google_snapped
```


## Sample trips snaps pipeline
```{r}
stops_sf <- stops %>% 
  st_as_sf(coords = c("stop_lon", "stop_lat"), crs = 4326, remove = FALSE) %>% 
  st_transform(26915)   # UTM zone 15N – works for Minneapolis

start_pts <- sample_trips %>% 
  st_as_sf(coords = c("start_lon", "start_lat"), crs = 4326, remove = FALSE) %>% 
  st_transform(26915)

start_nn_idx <- st_nearest_feature(start_pts, stops_sf)

start_snap <- stops_sf[start_nn_idx, c("stop_id","stop_lat","stop_lon")] %>% 
  st_drop_geometry() %>% 
  rename(
    start_stop_id_snap  = stop_id,
    start_stop_lat_snap = stop_lat,
    start_stop_lon_snap = stop_lon
  )

end_pts <- sample_trips %>% 
  st_as_sf(coords = c("end_lon", "end_lat"), crs = 4326, remove = FALSE) %>% 
  st_transform(26915)

end_nn_idx <- st_nearest_feature(end_pts, stops_sf)

end_snap <- stops_sf[end_nn_idx, c("stop_id","stop_lat","stop_lon")] %>% 
  st_drop_geometry() %>% 
  rename(
    end_stop_id_snap  = stop_id,
    end_stop_lat_snap = stop_lat,
    end_stop_lon_snap = stop_lon
  )

sample_trips_snap <- sample_trips %>% 
  bind_cols(start_snap, end_snap)

sample_trips_snap %>% 
  select(trip_id,
         start_lat, start_lon,
         start_stop_id_snap, start_stop_lat_snap, start_stop_lon_snap,
         end_lat, end_lon,
         end_stop_id_snap, end_stop_lat_snap, end_stop_lon_snap) %>% 
  head()
```

```{r}
stops_sf <- stops %>%
  st_as_sf(
    coords = c("stop_lon", "stop_lat"),
    crs    = 4326,
    remove = FALSE 
  )

start_pts <- sample_trips %>%
  transmute(trip_id,
            lon = start_lon,
            lat = start_lat) %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326)

idx_start <- st_nearest_feature(start_pts, stops_sf)

start_snap <- tibble(
  trip_id             = sample_trips$trip_id,
  start_stop_id_snap  = stops_sf$stop_id[idx_start],
  start_stop_lat_snap = stops_sf$stop_lat[idx_start],
  start_stop_lon_snap = stops_sf$stop_lon[idx_start]
)

end_pts <- sample_trips %>%
  transmute(trip_id,
            lon = end_lon,
            lat = end_lat) %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326)

idx_end <- st_nearest_feature(end_pts, stops_sf)

end_snap <- tibble(
  trip_id           = sample_trips$trip_id,
  end_stop_id_snap  = stops_sf$stop_id[idx_end],
  end_stop_lat_snap = stops_sf$stop_lat[idx_end],
  end_stop_lon_snap = stops_sf$stop_lon[idx_end]
)

sample_trips_snapped <- sample_trips %>%
  left_join(start_snap, by = "trip_id") %>%
  left_join(end_snap,   by = "trip_id")

head(sample_trips_snapped)
```
# Bike
## Google routing helper for bike and transit
```{r}
google_route <- function(origin_lat, origin_lon,
                         dest_lat, dest_lon,
                         mode = c("bicycling", "transit"),
                         departure_time = NULL,
                         api_key = Sys.getenv("GOOGLE_MAPS_KEY"),
                         debug = FALSE) {
  
  mode <- match.arg(mode)
  
  base_url <- "https://maps.googleapis.com/maps/api/directions/json"
  
  params <- list(
    origin      = sprintf("%.6f,%.6f", origin_lat, origin_lon),
    destination = sprintf("%.6f,%.6f", dest_lat,  dest_lon),
    mode        = mode,
    key         = api_key
  )
  
  if (!is.null(departure_time)) {
    params$departure_time <- as.character(departure_time)
  }
  
  resp <- GET(base_url, query = params)
  
  if (debug) {
    message("DEBUG URL: ", resp$url)
  }
  
  if (http_error(resp)) {
    return(tibble(
      duration_min  = NA_real_,
      route_summary = NA_character_,
      status        = "HTTP_ERROR",
      error_message = paste0("HTTP ", status_code(resp))
    ))
  }
  
  content_list <- content(resp, as = "parsed", type = "application/json")
  api_status   <- content_list$status %||% NA_character_
  
  if (!identical(api_status, "OK")) {
    err_msg <- content_list$error_message %||% NA_character_
    return(tibble(
      duration_min  = NA_real_,
      route_summary = NA_character_,
      status        = api_status,
      error_message = err_msg
    ))
  }
  
  route  <- content_list$routes[[1]]
  leg    <- route$legs[[1]]
  
  duration_sec <- leg$duration$value
  duration_min <- duration_sec / 60
  
  summary_txt <- paste0(leg$start_address, " → ", leg$end_address)
  
  tibble(
    duration_min  = duration_min,
    route_summary = summary_txt,
    status        = api_status,
    error_message = NA_character_
  )
}
```

## Full pipeline: micromobility vs snapped-transit vs bike
```{r}
sample_trips_snapped$StartTime <- as.POSIXct(sample_trips_snapped$StartTime, tz = "America/Chicago")

sample_trips_snapped <- sample_trips_snapped %>%
  mutate(departure_unix = as.integer(StartTime))

bike_results <- sample_trips_snapped %>%
  mutate(
    res = pmap(
      list(start_stop_lat_snap,
           start_stop_lon_snap,
           end_stop_lat_snap,
           end_stop_lon_snap,
           departure_unix),
      ~ google_route(
        origin_lat     = ..1,
        origin_lon     = ..2,
        dest_lat       = ..3,
        dest_lon       = ..4,
        mode           = "bicycling",
        departure_time = ..5
      )
    )
  ) %>%
  select(trip_id, res) %>%
  unnest(res) %>%
  rename(
    bike_min           = duration_min,
    bike_route_summary = route_summary,
    bike_status        = status,
    bike_error         = error_message
  )

transit_results <- sample_trips_snapped %>%
  mutate(
    res = pmap(
      list(start_stop_lat_snap,
           start_stop_lon_snap,
           end_stop_lat_snap,
           end_stop_lon_snap,
           departure_unix),
      ~ google_route(
        origin_lat     = ..1,
        origin_lon     = ..2,
        dest_lat       = ..3,
        dest_lon       = ..4,
        mode           = "transit",
        departure_time = ..5
      )
    )
  ) %>%
  select(trip_id, res) %>%
  unnest(res) %>%
  rename(
    transit_min           = duration_min,
    transit_route_summary = route_summary,
    transit_status        = status,
    transit_error         = error_message
  )
```

```{r}
google_results <- sample_trips_snapped %>%
  select(trip_id, duration_micro_min, StartTime,
         starts_with("start_stop"), starts_with("end_stop")) %>%
  left_join(bike_results,    by = "trip_id") %>%
  left_join(transit_results, by = "trip_id")
```

### Filter to only successful calls
```{r}
google_success <- google_results %>%
  filter(bike_status == "OK", transit_status == "OK")
```

```{r}
google_results %>%
  count(bike_status, transit_status)
```

### More debug
```{r}
sample_trips_snapped %>%
  count(trip_id) %>%
  filter(n > 1)

bike_results %>%
  count(trip_id) %>%
  filter(n > 1)

transit_results %>%
  count(trip_id) %>%
  filter(n > 1)
```

```{r}
bike_results_clean <- bike_results %>%
  arrange(trip_id) %>%
  group_by(trip_id) %>%
  slice(1) %>%
  ungroup()

transit_results_clean <- transit_results %>%
  arrange(trip_id) %>%
  group_by(trip_id) %>%
  slice(1) %>%
  ungroup()

saveRDS(bike_results_clean,    "data/bike_results_clean.rds")
saveRDS(transit_results_clean, "data/transit_results_clean.rds")
```

```{r}
bike_results_clean %>%
  count(trip_id) %>%
  filter(n > 1)

transit_results_clean %>%
  count(trip_id) %>%
  filter(n > 1)
```

```{r}
google_results <- sample_trips_snapped %>%
  select(
    trip_id,
    duration_micro_min,
    StartTime,
    starts_with("start_stop"),
    starts_with("end_stop")
  ) %>%
  left_join(
    bike_results_clean,
    by = "trip_id",
    relationship = "many-to-one"
  ) %>%
  left_join(
    transit_results_clean,
    by = "trip_id",
    relationship = "many-to-one"
  )
```

```{r}
nrow(sample_trips_snapped)
nrow(google_results)
```

```{r}
google_results %>%
  count(bike_status, transit_status)
```

## Re-do function for bike only
```{r}
# get_bike_duration <- function(origin_lat, origin_lon,
#                               dest_lat, dest_lon,
#                               api_key = Sys.getenv("GOOGLE_MAPS_KEY")) {
# 
#   base_url <- "https://maps.googleapis.com/maps/api/directions/json"
# 
#   query <- list(
#     origin      = sprintf("%.6f,%.6f", origin_lat, origin_lon),
#     destination = sprintf("%.6f,%.6f", dest_lat,  dest_lon),
#     mode        = "bicycling",
#     key         = api_key
#   )
# 
#   res <- GET(base_url, query = query)
#   txt <- content(res, as = "text", encoding = "UTF-8")
#   j   <- fromJSON(txt, flatten = TRUE)
# 
#   status <- j$status
# 
#   if (status != "OK") {
#     tibble(
#       bike_duration_min = NA_real_,
#       bike_status       = status,
#       bike_error        = if (!is.null(j$error_message)) j$error_message else NA_character_
#     )
#   } else {
#     dur_sec <- j$routes[[1]]$legs[[1]]$duration$value
#     tibble(
#       bike_duration_min = dur_sec / 60,
#       bike_status       = status,
#       bike_error        = NA_character_
#     )
#   }
# }
```

```{r}
get_bike_duration <- function(origin_lat, origin_lon,
                              dest_lat, dest_lon,
                              api_key = Sys.getenv("GOOGLE_MAPS_KEY")) {

  base_url <- "https://maps.googleapis.com/maps/api/directions/json"

  query <- list(
    origin      = sprintf("%.6f,%.6f", origin_lat, origin_lon),
    destination = sprintf("%.6f,%.6f", dest_lat,  dest_lon),
    mode        = "bicycling",
    key         = api_key
  )

  res <- GET(base_url, query = query)
  txt <- content(res, as = "text", encoding = "UTF-8")

  j   <- fromJSON(txt)

  status <- j$status %||% NA_character_

  if (status != "OK" ||
      is.null(j$routes) || length(j$routes) == 0 ||
      is.null(j$routes[[1]]$legs) || length(j$routes[[1]]$legs) == 0) {

    return(tibble(
      bike_duration_min = NA_real_,
      bike_status       = status,
      bike_error        = if (!is.null(j$error_message)) j$error_message else NA_character_
    ))
  }

  leg <- j$routes[[1]]$legs[[1]]
  dur_sec <- leg$duration$value

  tibble(
    bike_duration_min = dur_sec / 60,
    bike_status       = status,
    bike_error        = NA_character_
  )
}
```


### Test on a single row
```{r}
test_row <- sample_trips_snapped %>% slice(1)

test_bike <- get_bike_duration(
  origin_lat = test_row$start_stop_lat_snap,
  origin_lon = test_row$start_stop_lon_snap,
  dest_lat   = test_row$end_stop_lat_snap,
  dest_lon   = test_row$end_stop_lon_snap
)

test_bike
```

```{r}
test_row <- sample_trips_snapped %>% slice(1)

txt <- GET(
  "https://maps.googleapis.com/maps/api/directions/json",
  query = list(
    origin      = sprintf("%.6f,%.6f", test_row$start_stop_lat_snap, test_row$start_stop_lon_snap),
    destination = sprintf("%.6f,%.6f", test_row$end_stop_lat_snap,   test_row$end_stop_lon_snap),
    mode        = "bicycling",
    key         = Sys.getenv("GOOGLE_MAPS_KEY")
  )
) |> content(as = "text", encoding = "UTF-8")

j <- fromJSON(txt, flatten = TRUE)

names(j)
names(j$routes)
str(j$routes$legs[[1]])
```

```{r}
get_bike_duration <- function(origin_lat, origin_lon,
                              dest_lat, dest_lon,
                              api_key = Sys.getenv("GOOGLE_MAPS_KEY")) {

  base_url <- "https://maps.googleapis.com/maps/api/directions/json"

  query <- list(
    origin      = sprintf("%.6f,%.6f", origin_lat, origin_lon),
    destination = sprintf("%.6f,%.6f", dest_lat,  dest_lon),
    mode        = "bicycling",
    key         = api_key
  )

  res <- GET(base_url, query = query)
  txt <- content(res, as = "text", encoding = "UTF-8")

  j <- fromJSON(txt, flatten = TRUE)

  status <- j$status %||% NA_character_

  if (status != "OK" || is.null(j$routes) || nrow(j$routes) == 0) {
    return(tibble(
      bike_duration_min = NA_real_,
      bike_status       = status,
      bike_error        = if (!is.null(j$error_message)) j$error_message else NA_character_
    ))
  }

  legs_df <- j$routes$legs[[1]]

  if (is.null(legs_df) || nrow(legs_df) == 0 ||
      is.null(legs_df$duration.value)) {
    return(tibble(
      bike_duration_min = NA_real_,
      bike_status       = status,
      bike_error        = "No duration.value in legs"
    ))
  }

  dur_sec <- legs_df$duration.value[1]

  tibble(
    bike_duration_min = dur_sec / 60,
    bike_status       = status,
    bike_error        = NA_character_
  )
}
```

Retest
```{r}
test_row <- sample_trips_snapped %>% slice(1)

test_bike <- get_bike_duration(
  origin_lat = test_row$start_stop_lat_snap,
  origin_lon = test_row$start_stop_lon_snap,
  dest_lat   = test_row$end_stop_lat_snap,
  dest_lon   = test_row$end_stop_lon_snap
)

test_bike
```

IT FINALLY WORKS

# Scale to 100 trips
```{r}
rm(google_results, bike_results)
```

```{r}
sample_100 <- sample_trips_snapped %>% 
  slice(1:100)

get_bike_for_row <- function(start_lat, start_lon, end_lat, end_lon) {
  get_bike_duration(
    origin_lat = start_lat,
    origin_lon = start_lon,
    dest_lat   = end_lat,
    dest_lon   = end_lon
  )
}

bike_results <- sample_100 %>% 
  mutate(
    bike_res = pmap(
      list(start_stop_lat_snap, start_stop_lon_snap,
           end_stop_lat_snap, end_stop_lon_snap),
      get_bike_for_row
    )
  ) %>% 
  unnest(bike_res)
```

```{r}
bike_results %>% 
  count(bike_status, useNA = "ifany")
```


## Build google_results again
```{r}
google_results <- sample_100 %>% 
  select(trip_id, duration_micro_min, StartTime,
         starts_with("start_stop"), starts_with("end_stop")) %>% 
  left_join(
    bike_results %>% 
      select(trip_id, bike_duration_min, bike_status, bike_error),
    by = "trip_id"
  )
```
```{r}
google_results %>% 
  count(bike_status, useNA = "ifany")
```

```{r}
google_results %>%
  select(trip_id, duration_micro_min, bike_duration_min, bike_status) %>%
  head()
```

## Clean bike summary (per trip)
```{r}
bike_summary <- bike_results %>%
  filter(bike_status == "OK") %>%
  semi_join(sample_100, by = "trip_id") %>%
  group_by(trip_id) %>%
  summarise(
    bike_duration_min = min(bike_duration_min, na.rm = TRUE),
    bike_n_routes     = n(),
    .groups = "drop"
  )
```

```{r}
head(bike_summary)
```


## Clean transit summary
```{r}
names(transit_results)

transit_results <- transit_results %>% 
  rename(transit_duration_min = transit_min)
```

```{r}
transit_summary <- transit_results %>% 
  filter(transit_status == "OK") %>% 
  semi_join(sample_100, by = "trip_id") %>% 
  group_by(trip_id) %>% 
  summarise(
    transit_duration_min = min(transit_duration_min, na.rm = TRUE),
    transit_n_routes     = n(),
    .groups = "drop"
  )
```

```{r}
head(transit_summary)
```
## Final combined results
```{r}
comparison_100 <- sample_100 %>% 
  select(trip_id, duration_micro_min) %>% 
  left_join(bike_summary,    by = "trip_id") %>% 
  left_join(transit_summary, by = "trip_id")

head(comparison_100)
```

# Transit

```{r}
transit_results %>% 
  count(transit_status, transit_error, useNA = "ifany")
```

```{r}
debug_transit_call <- function(origin_lat, origin_lon,
                               dest_lat, dest_lon,
                               departure_time = NULL,
                               api_key = Sys.getenv("GOOGLE_MAPS_KEY")) {

  base_url <- "https://maps.googleapis.com/maps/api/directions/json"

  params <- list(
    origin      = sprintf("%.6f,%.6f", origin_lat, origin_lon),
    destination = sprintf("%.6f,%.6f", dest_lat,  dest_lon),
    mode        = "transit",
    key         = api_key
  )

  if (!is.null(departure_time)) {
    params$departure_time <- as.integer(departure_time)
  }

  resp <- httr::GET(base_url, query = params)

  cat("DEBUG URL:\n", resp$url, "\n\n")

  txt <- httr::content(resp, as = "text", encoding = "UTF-8")
  j   <- jsonlite::fromJSON(txt)

  str(list(
    status              = j$status,
    error_message       = j$error_message %||% NA_character_,
    available_modes     = j$available_travel_modes %||% NA_character_,
    n_routes            = length(j$routes)
  ))

  invisible(j)
}

test_row <- sample_trips_snapped %>% slice(1)

j_test <- debug_transit_call(
  origin_lat = test_row$start_stop_lat_snap,
  origin_lon = test_row$start_stop_lon_snap,
  dest_lat   = test_row$end_stop_lat_snap,
  dest_lon   = test_row$end_stop_lon_snap,
  departure_time = NULL
)
```

### Transit helper function
```{r}
get_transit_duration <- function(origin_lat, origin_lon,
                                 dest_lat, dest_lon,
                                 departure_time = NULL,
                                 api_key = Sys.getenv("GOOGLE_MAPS_KEY")) {

  base_url <- "https://maps.googleapis.com/maps/api/directions/json"

  params <- list(
    origin      = sprintf("%.6f,%.6f", origin_lat, origin_lon),
    destination = sprintf("%.6f,%.6f", dest_lat,  dest_lon),
    mode        = "transit",
    key         = api_key
  )

  if (!is.null(departure_time)) {
    params$departure_time <- as.integer(departure_time)
  }

  res <- httr::GET(base_url, query = params)
  txt <- httr::content(res, as = "text", encoding = "UTF-8")
  j   <- jsonlite::fromJSON(txt, flatten = TRUE)

  status <- j$status %||% NA_character_

  if (status != "OK" || is.null(j$routes) || nrow(j$routes) == 0) {
    return(tibble::tibble(
      transit_duration_min = NA_real_,
      transit_status       = status,
      transit_error        = j$error_message %||% NA_character_
    ))
  }

  legs_df <- j$routes$legs[[1]]

  if (is.null(legs_df) || nrow(legs_df) == 0 || is.null(legs_df$duration.value)) {
    return(tibble::tibble(
      transit_duration_min = NA_real_,
      transit_status       = status,
      transit_error        = "No duration.value in legs"
    ))
  }

  dur_sec <- legs_df$duration.value[1]

  tibble::tibble(
    transit_duration_min = dur_sec / 60,
    transit_status       = status,
    transit_error        = NA_character_
  )
}
```

### Test on single row
```{r}
test_transit <- get_transit_duration(
  origin_lat = test_row$start_stop_lat_snap,
  origin_lon = test_row$start_stop_lon_snap,
  dest_lat   = test_row$end_stop_lat_snap,
  dest_lon   = test_row$end_stop_lon_snap,
  departure_time = as.integer(Sys.time())
)

test_transit
```

### Scale to 100 trips
```{r}
transit_results <- sample_100 %>%
  mutate(
    transit_res = purrr::pmap(
      list(start_stop_lat_snap,
           start_stop_lon_snap,
           end_stop_lat_snap,
           end_stop_lon_snap,
           departure_unix),
      ~ get_transit_duration(..1, ..2, ..3, ..4, departure_time = ..5)
    )
  ) %>%
  tidyr::unnest(transit_res)
```

```{r}
transit_results %>%
  count(transit_status, useNA = "ifany")
```




```{r}
google_route0 <- function(origin_lat, origin_lon,
                         dest_lat, dest_lon,
                         mode = c("bicycling", "transit"),
                         departure_time = NULL,
                         api_key = Sys.getenv("GOOGLE_MAPS_KEY")) {
  
  mode <- match.arg(mode)

  base_url <- "https://maps.googleapis.com/maps/api/directions/json"

  params <- list(
    origin      = sprintf("%.6f,%.6f", origin_lat, origin_lon),
    destination = sprintf("%.6f,%.6f", dest_lat,  dest_lon),
    mode        = mode,
    key         = api_key
  )

  if (!is.null(departure_time)) {
    params$departure_time <- as.integer(departure_time)
  }

  res  <- GET(base_url, query = params)
  txt  <- content(res, "text", encoding = "UTF-8")
  j    <- fromJSON(txt, flatten = TRUE)

  status <- j$status

  if (status != "OK") {
    return(tibble(
      duration_min = NA_real_,
      status = status,
      error_message = j$error_message %||% NA_character_
    ))
  }

  leg <- j$routes$legs[[1]]

  tibble(
    duration_min  = leg$duration.value / 60,
    status        = status,
    error_message = NA_character_
  )
}
```

```{r}
test <- google_route0(
  origin_lat  = 44.9738,
  origin_lon  = -93.2270,
  dest_lat    = 44.9790,
  dest_lon    = -93.2658,
  mode        = "transit",
  departure_time = as.integer(Sys.time())
)

test
```

```{r}
sample_trips_snapped0 <- sample_trips_snapped %>%
  mutate(departure_unix = as.integer(StartTime))
```


```{r}
transit_results <- sample_trips_snapped %>%
  mutate(
    res = pmap(
      list(start_stop_lat_snap,
           start_stop_lon_snap,
           end_stop_lat_snap,
           end_stop_lon_snap),
      ~ google_route(
        origin_lat     = ..1,
        origin_lon     = ..2,
        dest_lat       = ..3,
        dest_lon       = ..4,
        mode           = "transit",
        departure_time = NULL,
        debug          = FALSE
      )
    )
  ) %>%
  select(trip_id, res) %>%
  unnest(res) %>%
  rename(
    transit_min           = duration_min,
    transit_route_summary = route_summary,
    transit_status        = status,
    transit_error         = error_message
  )
```

```{r}
transit_results %>%
  count(transit_status, useNA = "ifany")
```

```{r}
r1 <- sample_trips_snapped %>% slice(1)

google_route(
  origin_lat  = r1$start_stop_lat_snap,
  origin_lon  = r1$start_stop_lon_snap,
  dest_lat    = r1$end_stop_lat_snap,
  dest_lon    = r1$end_stop_lon_snap,
  mode        = "transit",
  departure_time = NULL,
  debug = TRUE
)
```

### final result for transit?
```{r}
transit_results0 <- sample_trips_snapped0 %>%
  mutate(
    res = pmap(
      list(start_stop_lat_snap,
           start_stop_lon_snap,
           end_stop_lat_snap,
           end_stop_lon_snap),
      ~ google_route(
        origin_lat     = ..1,
        origin_lon     = ..2,
        dest_lat       = ..3,
        dest_lon       = ..4,
        mode           = "transit",
        departure_time = NULL,
        debug          = FALSE
      )
    )
  ) %>%
  select(trip_id, res) %>%
  unnest(res) %>%
  rename(
    transit_duration_min = duration_min,
    transit_route_summary = route_summary,
    transit_status        = status,
    transit_error         = error_message
  )
```

check
```{r}
transit_results %>%
  count(transit_status, useNA = "ifany")

transit_results <- transit_results %>% 
  rename(
    transit_duration_min = transit_min
  )
```

### clean transit_summary (one row per trip)
```{r}
transit_summary <- transit_results %>% 
  filter(transit_status == "OK") %>%
  semi_join(sample_100, by = "trip_id") %>%
  group_by(trip_id) %>% 
  summarise(
    transit_duration_min = min(transit_duration_min, na.rm = TRUE),
    transit_n_routes     = n(),
    .groups = "drop"
  )
```

```{r}
head(transit_summary)
```

# Final combined results
```{r}
comparison_100 <- sample_100 %>%
  select(trip_id, duration_micro_min) %>%
  left_join(bike_summary,    by = "trip_id") %>%
  left_join(transit_summary, by = "trip_id")
```

```{r}
comparison_100 %>%
  select(trip_id, duration_micro_min,
         bike_duration_min, transit_duration_min) %>%
  head()
```

```{r}
comparison_100 <- comparison_100 %>%
  mutate(
    diff_bike_micro    = bike_duration_min    - duration_micro_min,
    diff_transit_micro = transit_duration_min - duration_micro_min,
    diff_transit_bike  = transit_duration_min - bike_duration_min
  )
```















