---
title: "Google Directions: Micromobility vs. Transit"
author: "Hanna Chang"
format: html
editor: visual
---

---
title: "Google Directions: Micromobility vs. Transit"
author: "Hanna Chang"
format: html
editor: visual
---

```{r setup, warning=FALSE, message=FALSE}
library(httr)
library(jsonlite)
library(dplyr)
library(purrr)
library(tidyr)
library(sf)
library(tibble)
```

```{r load-data, warning=FALSE, message=FALSE}
#load("apienvironment.RData")
#stops <- read.csv("stops.csv")
#sample_trips <- read.csv("sample_trips.csv")
```

```{r api-key, warning=FALSE, message=FALSE}
Sys.setenv(GOOGLE_MAPS_KEY = "")

google_api_key <- Sys.getenv("GOOGLE_MAPS_KEY")
if (google_api_key == "") {
  stop("GOOGLE_MAPS_KEY is not set. Please set it via Sys.setenv().")
}
```

# Snap trip endpoints to nearest transit stops

Use a projected CRS for better distance handling (UTM zone 15N for Minneapolis)
```{r snap-to-stops, warning=FALSE, message=FALSE}
stops_sf <- stops %>%
st_as_sf(coords = c("stop_lon", "stop_lat"), crs = 4326, remove = FALSE) %>%
st_transform(26915)
```

Start points
```{r snap-to-stops1, warning=FALSE, message=FALSE}
start_pts <- sample_trips %>%
st_as_sf(coords = c("start_lon", "start_lat"), crs = 4326, remove = FALSE) %>%
st_transform(26915)

start_nn_idx <- st_nearest_feature(start_pts, stops_sf)

start_snap <- stops_sf[start_nn_idx, c("stop_id", "stop_lat", "stop_lon")] %>%
st_drop_geometry() %>%
rename(
start_stop_id_snap  = stop_id,
start_stop_lat_snap = stop_lat,
start_stop_lon_snap = stop_lon
)
```

End points
```{r snap-to-stops2, warning=FALSE, message=FALSE}
end_pts <- sample_trips %>%
st_as_sf(coords = c("end_lon", "end_lat"), crs = 4326, remove = FALSE) %>%
st_transform(26915)

end_nn_idx <- st_nearest_feature(end_pts, stops_sf)

end_snap <- stops_sf[end_nn_idx, c("stop_id", "stop_lat", "stop_lon")] %>%
st_drop_geometry() %>%
rename(
end_stop_id_snap  = stop_id,
end_stop_lat_snap = stop_lat,
end_stop_lon_snap = stop_lon
)
```

Combine snapped stops with original trip data
```{r snap-to-stops3, warning=FALSE, message=FALSE}
sample_trips_snapped <- sample_trips %>%
bind_cols(start_snap, end_snap)
```

Sanity check
```{r snap-to-stops4, warning=FALSE, message=FALSE}
head(
sample_trips_snapped %>%
select(
trip_id,
start_lat, start_lon,
start_stop_id_snap, start_stop_lat_snap, start_stop_lon_snap,
end_lat, end_lon,
end_stop_id_snap, end_stop_lat_snap, end_stop_lon_snap
)
)
```

# Google Directions helper functions

Small helper to safely pull an error message from the JSON
```{r helpers-google, warning=FALSE, message=FALSE}
get_error_message <- function(j) {
if (!is.null(j$error_message)) j$error_message else NA_character_
}
```

Bike duration helper
```{r bike-helper, warning=FALSE, message=FALSE}
get_bike_duration <- function(origin_lat, origin_lon,
                              dest_lat, dest_lon,
                              api_key = Sys.getenv("GOOGLE_MAPS_KEY"),
                              debug = FALSE) {

  api_key <- trimws(api_key)
  if (api_key == "") {
    return(tibble(
      bike_duration_min = NA_real_,
      bike_status       = "NO_API_KEY",
      bike_error        = "GOOGLE_MAPS_KEY not set"
    ))
  }

  coords <- c(origin_lat, origin_lon, dest_lat, dest_lon)
  if (any(!is.finite(coords))) {
    return(tibble(
      bike_duration_min = NA_real_,
      bike_status       = "INVALID_COORDS",
      bike_error        = paste("Non-finite coordinates:",
                                paste(coords, collapse = ", "))
    ))
  }

  base_url <- "https://maps.googleapis.com/maps/api/directions/json"

  query <- list(
    origin      = sprintf("%.6f,%.6f", origin_lat, origin_lon),
    destination = sprintf("%.6f,%.6f", dest_lat,  dest_lon),
    mode        = "bicycling",
    key         = api_key
  )

  full_url <- httr::modify_url(base_url, query = query)

  if (debug) {
    message("BIKE URL: ", full_url)
  }

  res <- httr::GET(full_url)

  if (httr::http_error(res)) {
    return(tibble(
      bike_duration_min = NA_real_,
      bike_status       = "HTTP_ERROR",
      bike_error        = paste0("HTTP ", httr::status_code(res))
    ))
  }

  txt <- httr::content(res, as = "text", encoding = "UTF-8")
  j   <- jsonlite::fromJSON(txt, flatten = TRUE)

  status <- j$status %||% NA_character_

  if (!identical(status, "OK") || is.null(j$routes) || nrow(j$routes) == 0) {
    return(tibble(
      bike_duration_min = NA_real_,
      bike_status       = status,
      bike_error        = if (!is.null(j$error_message)) j$error_message else NA_character_
    ))
  }

  legs_df <- j$routes$legs[[1]]

  if (is.null(legs_df) || nrow(legs_df) == 0 || is.null(legs_df$duration.value)) {
    return(tibble(
      bike_duration_min = NA_real_,
      bike_status       = status,
      bike_error        = "No duration.value in legs"
    ))
  }

  dur_sec <- legs_df$duration.value[1]

  tibble(
    bike_duration_min = dur_sec / 60,
    bike_status       = status,
    bike_error        = NA_character_
  )
}
```

Transit duration helper
```{r transit-helper, warning=FALSE, message=FALSE}
get_transit_duration <- function(origin_lat, origin_lon,
                                 dest_lat, dest_lon,
                                 departure_time = NULL,
                                 api_key = Sys.getenv("GOOGLE_MAPS_KEY"),
                                 debug = FALSE) {

  api_key <- trimws(api_key)
  if (api_key == "") {
    return(tibble(
      transit_duration_min = NA_real_,
      transit_status       = "NO_API_KEY",
      transit_error        = "GOOGLE_MAPS_KEY not set"
    ))
  }

  coords <- as.numeric(c(origin_lat, origin_lon, dest_lat, dest_lon))
  if (any(!is.finite(coords))) {
    return(tibble(
      transit_duration_min = NA_real_,
      transit_status       = "INVALID_COORDS",
      transit_error        = paste("Non-finite coordinates:",
                                   paste(coords, collapse = ", "))
    ))
  }

  base_url <- "https://maps.googleapis.com/maps/api/directions/json"

  params <- list(
    origin      = sprintf("%.6f,%.6f", coords[1], coords[2]),
    destination = sprintf("%.6f,%.6f", coords[3], coords[4]),
    mode        = "transit",
    key         = api_key
  )

  if (!is.null(departure_time)) {
    params$departure_time <- as.integer(departure_time)
  }

 run_call <- function(params) {
    full_url <- httr::modify_url(base_url, query = params)
    if (debug) message("TRANSIT URL: ", full_url)

    res <- httr::GET(full_url)
    if (httr::http_error(res)) {
      return(list(
        status = "HTTP_ERROR",
        error  = paste0("HTTP ", httr::status_code(res)),
        j      = NULL
      ))
    }
    txt <- httr::content(res, as = "text", encoding = "UTF-8")
    j   <- jsonlite::fromJSON(txt, flatten = TRUE)
    list(status = j$status %||% NA_character_, error = j$error_message %||% NA_character_, j = j)
  }

  call1 <- run_call(params)

  if (identical(call1$status, "ZERO_RESULTS")) {
    params2 <- params
    params2$departure_time <- NULL
    call2 <- run_call(params2)
    status <- call2$status
    error  <- call2$error
    j      <- call2$j
  } else {
    status <- call1$status
    error  <- call1$error
    j      <- call1$j
  }

  if (!identical(status, "OK") || is.null(j$routes) || nrow(j$routes) == 0) {
    return(tibble(
      transit_duration_min = NA_real_,
      transit_status       = status,
      transit_error        = error
    ))
  }

  legs_df <- j$routes$legs[[1]]
  if (is.null(legs_df) || nrow(legs_df) == 0 || is.null(legs_df$duration.value)) {
    return(tibble(
      transit_duration_min = NA_real_,
      transit_status       = status,
      transit_error        = "No duration.value in legs"
    ))
  }

  dur_sec <- legs_df$duration.value[1]

  tibble(
    transit_duration_min = dur_sec / 60,
    transit_status       = status,
    transit_error        = NA_character_
  )
}
```

# Prepare subset and departure times

```{r prepare-departure, warning=FALSE, message=FALSE}
sample_trips_snapped <- sample_trips_snapped %>%
mutate(
StartTime      = as.POSIXct(StartTime, tz = "America/Chicago"),
departure_unix = as.integer(StartTime)
)
```

Number of trips to call the API on
```{r sample-n-trips, warning=FALSE, message=FALSE}
N_TRIPS <- 100

sample_N <- sample_trips_snapped %>%
slice(1:N_TRIPS)
```

# Bike durations for N trips

```{r bike-results, warning=FALSE, message=FALSE}
get_bike_for_row <- function(start_lat, start_lon, end_lat, end_lon) {
get_bike_duration(
origin_lat = start_lat,
origin_lon = start_lon,
dest_lat   = end_lat,
dest_lon   = end_lon
)
}

bike_results <- sample_N %>%
mutate(
bike_res = pmap(
list(
start_stop_lat_snap,
start_stop_lon_snap,
end_stop_lat_snap,
end_stop_lon_snap
),
get_bike_for_row
)
) %>%
unnest(bike_res)
```

Status check
```{r bike-status-check, warning=FALSE, message=FALSE}
bike_results %>%
count(bike_status, useNA = "ifany")
```


### Test a single row

```{r bike-test-single, warning=FALSE, message=FALSE}
r1 <- sample_N %>% slice(1)

test_bike <- get_bike_duration(
  origin_lat = r1$start_stop_lat_snap,
  origin_lon = r1$start_stop_lon_snap,
  dest_lat   = r1$end_stop_lat_snap,
  dest_lon   = r1$end_stop_lon_snap,
  debug      = TRUE
)

test_bike
```

### Rebuild bike_results with updated function

```{r bike-results-updated, warning=FALSE, message=FALSE}
bike_results <- sample_N %>%
  mutate(
    bike_res = purrr::pmap(
      list(
        start_stop_lat_snap,
        start_stop_lon_snap,
        end_stop_lat_snap,
        end_stop_lon_snap
      ),
      ~ get_bike_duration(..1, ..2, ..3, ..4)
    )
  ) %>%
  tidyr::unnest(bike_res)
```


# Transit durations for N trips

```{r transit-results, warning=FALSE, message=FALSE}
get_transit_for_row <- function(start_lat, start_lon, end_lat, end_lon, departure_unix) {
get_transit_duration(
origin_lat      = start_lat,
origin_lon      = start_lon,
dest_lat        = end_lat,
dest_lon        = end_lon,
departure_time  = departure_unix
)
}

transit_results <- sample_N %>%
mutate(
transit_res = pmap(
list(
start_stop_lat_snap,
start_stop_lon_snap,
end_stop_lat_snap,
end_stop_lon_snap,
departure_unix
),
get_transit_for_row
)
) %>%
unnest(transit_res)
```

Status check
```{r transit-status-check, warning=FALSE, message=FALSE}
transit_results %>%
count(transit_status, useNA = "ifany")
```

```{r transit-test-single, warning=FALSE, message=FALSE}
r1 <- sample_N %>% slice(1)

test_transit <- get_transit_duration(
  origin_lat     = r1$start_stop_lat_snap,
  origin_lon     = r1$start_stop_lon_snap,
  dest_lat       = r1$end_stop_lat_snap,
  dest_lon       = r1$end_stop_lon_snap,
  departure_time = r1$departure_unix,
  debug          = TRUE
)

test_transit
```


# Summaries per trip

```{r bike-summary, warning=FALSE, message=FALSE}
bike_summary <- bike_results %>%
filter(bike_status == "OK") %>%
group_by(trip_id) %>%
summarise(
bike_duration_min = if (all(is.na(bike_duration_min))) {
NA_real_
} else {
min(bike_duration_min, na.rm = TRUE)
},
bike_n_routes = sum(!is.na(bike_duration_min)),
.groups = "drop"
)

head(bike_summary)
```

```{r transit-summary, warning=FALSE, message=FALSE}
transit_summary <- transit_results %>%
filter(transit_status == "OK") %>%
group_by(trip_id) %>%
summarise(
transit_duration_min = if (all(is.na(transit_duration_min))) {
NA_real_
} else {
min(transit_duration_min, na.rm = TRUE)
},
transit_n_routes = sum(!is.na(transit_duration_min)),
.groups = "drop"
)

head(transit_summary)
```

# Final combined comparison

```{r comparison, warning=FALSE, message=FALSE}
#| tbl-cap: "Comparison of micromobility, bike, and transit durations (first few trips)."

comparison_N <- sample_N %>%
select(trip_id, duration_micro_min) %>%
left_join(bike_summary,    by = "trip_id") %>%
left_join(transit_summary, by = "trip_id") %>%
mutate(
diff_bike_micro    = bike_duration_min    - duration_micro_min,
diff_transit_micro = transit_duration_min - duration_micro_min,
diff_transit_bike  = transit_duration_min - bike_duration_min
)

comparison_N %>%
select(
trip_id,
duration_micro_min,
bike_duration_min,
transit_duration_min,
diff_bike_micro,
diff_transit_micro,
diff_transit_bike
) %>%
head()
```


```{r}
google_results_clean <- google_results %>%
transmute(
    trip_id,

    start_lat, start_lon,
    end_lat, end_lon,

    StartTime = StartTime.x,

    start_stop_id_snap = start_stop_id_snap.x,
    start_stop_lat_snap = start_stop_lat_snap.x,
    start_stop_lon_snap = start_stop_lon_snap.x,

    end_stop_id_snap = end_stop_id_snap.x,
    end_stop_lat_snap = end_stop_lat_snap.x,
    end_stop_lon_snap = end_stop_lon_snap.x,

    duration_micro_min = duration_micro_min.x,
    bike_duration_min,
    transit_duration_min
  )
```

```{r}
google_results_long <- google_results_clean %>%
  pivot_longer(
    cols = c(duration_micro_min, bike_duration_min, transit_duration_min),
    names_to = "mode_raw",
    values_to = "duration_min"
  ) %>%
  mutate(
    mode = recode(
      mode_raw,
      duration_micro_min = "scooter",
      bike_duration_min = "bike",
      transit_duration_min = "transit"
    )
  ) %>%
  select(
    trip_id, StartTime,
    start_lat, start_lon, end_lat, end_lon,
    start_stop_id_snap, start_stop_lat_snap, start_stop_lon_snap,
    end_stop_id_snap, end_stop_lat_snap, end_stop_lon_snap,
    mode, duration_min
  )
```

```{r}
google_results_long <- google_results %>%
transmute(
    trip_id,
    StartTime = StartTime.x,

    start_lat, start_lon,
    end_lat, end_lon,

    start_stop_id_snap = start_stop_id_snap.x,
    start_stop_lat_snap = start_stop_lat_snap.x,
    start_stop_lon_snap = start_stop_lon_snap.x,

    end_stop_id_snap   = end_stop_id_snap.x,
    end_stop_lat_snap  = end_stop_lat_snap.x,
    end_stop_lon_snap  = end_stop_lon_snap.x,

    duration_micro_min = duration_micro_min.x,
    bike_duration_min,
    transit_duration_min
  ) %>%
pivot_longer(
    cols = c(duration_micro_min, bike_duration_min, transit_duration_min),
    names_to = "mode_raw",
    values_to = "duration_min"
  ) %>%
mutate(
    mode = recode(
      mode_raw,
      "duration_micro_min" = "scooter",
      "bike_duration_min" = "bike",
      "transit_duration_min" = "transit"
    )
  ) %>%
select(
    trip_id, StartTime,
    start_lat, start_lon, end_lat, end_lon,
    start_stop_id_snap, start_stop_lat_snap, start_stop_lon_snap,
    end_stop_id_snap, end_stop_lat_snap, end_stop_lon_snap,
    mode, duration_min
  )
```



```{r save-google-results}
saveRDS(bike_results,         "data/bike_results.rds")
saveRDS(bike_summary,         "data/bike_summary.rds")
saveRDS(transit_results,      "data/transit_results.rds")
saveRDS(transit_summary,         "data/transit_summary.rds")
saveRDS(comparison_N,         "data/comparison_N.rds")
saveRDS(google_results_clean, "data/google_results_wide.rds")
saveRDS(google_results_long,  "data/google_results_long.rds")
```

